---
title: "R Notebook"
output: html_notebook
---


The Dogs vs. Cats dataset that you’ll use isn’t packaged with Keras. It was made available by Kaggle as part of a computer-vision competition in late 2013, back when convnets weren’t mainstream. You can download the original dataset from www.kaggle.com/c/dogs-vs-cats/data 

***
Copying images to train, validation, and test directories
***

```{r}
original_dataset_dir <- "~/datasets/catsndogs/train"

base_dir <- "~/Downloads/cats_and_dogs_small" 
dir.create(base_dir)
train_dir <- file.path(base_dir, "train") 
dir.create(train_dir)

validation_dir <- file.path(base_dir, "validation") 
dir.create(validation_dir)
test_dir <- file.path(base_dir, "test") 
dir.create(test_dir)
train_cats_dir <- file.path(train_dir, "cats") 
dir.create(train_cats_dir)

train_dogs_dir <- file.path(train_dir, "dogs") 
dir.create(train_dogs_dir)

validation_cats_dir <- file.path(validation_dir, "cats") 
dir.create(validation_cats_dir)

validation_dogs_dir <- file.path(validation_dir, "dogs") 
dir.create(validation_dogs_dir)

test_cats_dir <- file.path(test_dir, "cats")
dir.create(test_cats_dir)
test_dogs_dir <- file.path(test_dir, "dogs") 
dir.create(test_dogs_dir)

fnames <- paste0("cat.", 1:1000, ".jpg") 
file.copy(file.path(original_dataset_dir, fnames),file.path(train_cats_dir))

fnames <- paste0("cat.", 1001:1500, ".jpg") 
file.copy(file.path(original_dataset_dir, fnames),
file.path(validation_cats_dir))

fnames <- paste0("cat.", 1501:2000, ".jpg")
file.copy(file.path(original_dataset_dir, fnames),
file.path(test_cats_dir))

fnames <- paste0("dog.", 1:1000, ".jpg") 
file.copy(file.path(original_dataset_dir, fnames),
file.path(train_dogs_dir))

fnames <- paste0("dog.", 1001:1500, ".jpg") 
file.copy(file.path(original_dataset_dir, fnames),
file.path(validation_dogs_dir))

fnames <- paste0("dog.", 1501:2000, ".jpg") 
file.copy(file.path(original_dataset_dir, fnames),
file.path(test_dogs_dir))

# As a sanity check, let’s count how many pictures are in each training split (train/validation/test):

cat("total training cat images:", length(list.files(train_cats_dir)), "\n")
cat("total training dog images:", length(list.files(train_dogs_dir)), "\n")
cat("total validation cat images:", length(list.files(validation_cats_dir)), "\n")
cat("total validation dog images:", length(list.files(validation_dogs_dir)), "\n")
cat("total test cat images:", length(list.files(test_cats_dir)), "\n")
cat("total test dog images:", length(list.files(test_dogs_dir)), "\n")
```

***
Instantiating a small convnet for cats vs. dogs classification
***

```{r}
library(keras)
model <- keras_model_sequential() %>%
layer_conv_2d(filters = 32, kernel_size = c(3, 3), activation = "relu",
                input_shape = c(150, 150, 3)) %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 64, kernel_size = c(3, 3), activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 128, kernel_size = c(3, 3), activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 128, kernel_size = c(3, 3), activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
   layer_flatten() %>%
  layer_dense(units = 512, activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")
```

```{r}
summary(model)
```

***
Configuring the model for training
***

```{r}
model %>% compile(
loss = "binary_crossentropy",
optimizer = optimizer_rmsprop(lr = 1e-4), metrics = c("acc")
)
```

***
Data Preprocessing

Currently, the data sits on a drive as JPEG files, so the steps for getting it into the network are roughly as follows:

1. Read the picture files.
2. Decode the JPEG content to RGB grids of pixels. 
3. Convert these into floating-point tensors.
4. Rescale the pixel values (between 0 and 255) to the [0, 1] interval (as you know, neural networks prefer to deal with small input values).

It may seem a bit daunting, but thankfully Keras has utilities to take care of these steps automatically. Keras includes a number of image processing helper tools. In particular, it includes the image_data_generator() function, which can automatically turn image files on disk into batches of pre-processed tensors. This is what we will use here.

***

```{r}
train_datagen <- image_data_generator(rescale = 1/255) 
validation_datagen <- image_data_generator(rescale = 1/255)

train_generator <- flow_images_from_directory( train_dir,
train_datagen,
target_size = c(150, 150),
  batch_size = 20,
  class_mode = "binary"
)
validation_generator <- flow_images_from_directory( validation_dir,
validation_datagen,
target_size = c(150, 150),
  batch_size = 20,
  class_mode = "binary"
)
```

***
Display a batch of data and labels
*** 

```{r}
batch <- generator_next(train_generator)
str(batch)
```

***
Fitting the model using a batch generator
***

```{r}
history <- model %>% fit_generator( train_generator,
steps_per_epoch = 100,
epochs = 30,
  validation_data = validation_generator,
  validation_steps = 50
)
```


***
Save the model
***

```{r}
 model %>% save_model_hdf5("cats_and_dogs_small_1.h5")
```


```{r}
plot(history)
```


***
Data augmentation
***

```{r}
datagen <- image_data_generator( 
  rescale = 1/255, 
  rotation_range = 40,
  width_shift_range = 0.2,
  height_shift_range = 0.2,
  shear_range = 0.2,
  zoom_range = 0.2, 
  horizontal_flip = TRUE, 
  fill_mode = "nearest"
)
```

```{r}
fnames <- list.files(train_cats_dir, full.names = TRUE) 
img_path <- fnames[[3]]
img <- image_load(img_path, target_size = c(150, 150)) 
img_array <- image_to_array(img)
img_array <- array_reshape(img_array, c(1, 150, 150, 3))
augmentation_generator <- flow_images_from_data( 
  img_array,
  generator = datagen,
  batch_size = 1
)
op <- par(mfrow = c(2, 2), pty = "s", mar = c(1, 0, 1, 0)) 
for (i in 1:4) {
batch <- generator_next(augmentation_generator)
plot(as.raster(batch[1,,,])) }
par(op)
```

